\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2020-03-06 analysis/paper/paper.tex
Fixed lollipop plot
to do:
fix casebase absoluterisk for glmnet. 
fix ??? tags

\begin{CodeInput}
R> data(ERSPC)
R> ERSPC$ScrArm <- factor(ERSPC$ScrArm, 
R>                        levels = c(0, 1), 
R>                        labels = c("Control group", "Screening group"))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> pt_object_strat <- casebase::popTime(ERSPC, 
R>                                      event = "DeadOfPrCa", 
R>                                      exposure = "ScrArm")
R> plot(pt_object_strat)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> cox_model <- survival::coxph(Surv(Follow.Up.Time, DeadOfPrCa) ~ ScrArm, 
R>                              data = ERSPC)
R> summary(cox_model)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> new_data <- data.frame(ScrArm = c("Control group", "Screening group"),
R>                        ignore = 99)
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>      xlab = "Years since Randomization", 
R>      ylab = "Cumulative Incidence (%)", 
R>      fun = "event",
R>      xlim = c(0,15), conf.int = FALSE, col = c("red","blue"), 
R>      main = sprintf("Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>                     exp(coef(cox_model)), 
R>                     exp(confint(cox_model))[1], 
R>                     exp(confint(cox_model))[2]),
R>      ylim = c(0, 2), yscale = 100)
R> legend("topleft", 
R>        legend = c("Control group", "Screening group"), 
R>        col = c("red","blue"),
R>        lty = c(1, 1), 
R>        bg = "gray90")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> casebase_exponential <- casebase::fitSmoothHazard(DeadOfPrCa ~ ScrArm, 
R>                                                   data = ERSPC, 
R>                                                   ratio = 100)
R> 
R> summary(casebase_exponential)
R> exp(coef(casebase_exponential)[2])
R> exp(confint(casebase_exponential)[2,])
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> smooth_risk_exp <- casebase::absoluteRisk(object = casebase_exponential, 
R>                                           time = seq(0,15,0.1), 
R>                                           newdata = new_data)
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>      xlab = "Years since Randomization", 
R>      ylab = "Cumulative Incidence (%)", 
R>      fun = "event",
R>      xlim = c(0,15), conf.int = FALSE, col = c("red","blue"), 
R>      main = sprintf("Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>                     exp(coef(cox_model)), 
R>                     exp(confint(cox_model))[1], 
R>                     exp(confint(cox_model))[2]),
R>      ylim = c(0, 2), yscale = 100)
R> lines(smooth_risk_exp[,1], smooth_risk_exp[,2], col = "red", lty = 2)
R> lines(smooth_risk_exp[,1], smooth_risk_exp[,3], col = "blue", lty = 2)
R> 
R> 
R> legend("topleft", 
R>        legend = c("Control group (Cox)","Control group (Casebase)",
R>                   "Screening group (Cox)", "Screening group (Casebase)"), 
R>        col = c("red","red", "blue","blue"),
R>        lty = c(1, 2, 1, 2), 
R>        bg = "gray90")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> casebase_time <- fitSmoothHazard(DeadOfPrCa ~ Follow.Up.Time + ScrArm, 
R>                                  data = ERSPC, 
R>                                  ratio = 100)
R> 
R> summary(casebase_time)
R> exp(coef(casebase_time))
R> exp(confint(casebase_time))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> smooth_risk_time <- casebase::absoluteRisk(object = casebase_time, 
R>                                           time = seq(0,15,0.1), 
R>                                           newdata = new_data)
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>      xlab = "Years since Randomization", 
R>      ylab = "Cumulative Incidence (%)", 
R>      fun = "event",
R>      xlim = c(0,15), conf.int = FALSE, col = c("red","blue"), 
R>      main = sprintf("Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>                     exp(coef(cox_model)), 
R>                     exp(confint(cox_model))[1], 
R>                     exp(confint(cox_model))[2]),
R>      ylim = c(0, 2), yscale = 100)
R> lines(smooth_risk_time[,1], smooth_risk_time[,2], col = "red", lty = 2)
R> lines(smooth_risk_time[,1], smooth_risk_time[,3], col = "blue", lty = 2)
R> 
R> legend("topleft", 
R>        legend = c("Control group (Cox)","Control group (Casebase)",
R>                   "Screening group (Cox)", "Screening group (Casebase)"), 
R>        col = c("red","red", "blue","blue"),
R>        lty = c(1, 2, 1, 2), 
R>        bg = "gray90")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> casebase_splines <- fitSmoothHazard(DeadOfPrCa ~ bs(Follow.Up.Time) + ScrArm, 
R>                                     data = ERSPC, 
R>                                     ratio = 100)
R> 
R> summary(casebase_splines)
R> exp(coef(casebase_splines))
R> exp(confint(casebase_splines))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> smooth_risk_splines <- absoluteRisk(object = casebase_splines, 
R>                                     time = seq(0,15,0.1), 
R>                                     newdata = new_data)
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>      xlab = "Years since Randomization", 
R>      ylab = "Cumulative Incidence (%)", 
R>      fun = "event",
R>      xlim = c(0,15), conf.int = FALSE, col = c("red","blue"), 
R>      main = sprintf("Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>                     exp(coef(cox_model)), 
R>                     exp(confint(cox_model))[1], 
R>                     exp(confint(cox_model))[2]),
R>      ylim = c(0, 2), yscale = 100)
R> lines(smooth_risk_splines[,1], smooth_risk_splines[,2], col = "red", lty = 2)
R> lines(smooth_risk_splines[,1], smooth_risk_splines[,3], col = "blue", lty = 2)
R> 
R> legend("topleft", 
R>        legend = c("Control group (Cox)","Control group (Casebase)",
R>                   "Screening group (Cox)", "Screening group (Casebase)"), 
R>        col = c("red","red", "blue","blue"),
R>        lty = c(1, 2, 1, 2), 
R>        bg = "gray90")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> c("Exp." = AIC(casebase_exponential),
R>   "Weib." = AIC(casebase_time),
R>   "Splines" = AIC(casebase_splines))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> model_cb <- fitSmoothHazard(
R>     Status ~ ftime + Sex + D + Phase + Source + Age, 
R>     data = bmtcrr, 
R>     ratio = 100, 
R>     time = "ftime")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(survival)
R> # Treat competing event as censoring
R> model_cox <- coxph(Surv(ftime, Status == 1) ~ Sex + D + Phase + Source + Age,
R>                    data = bmtcrr)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Pick 100 equidistant points between 0 and 60 months
R> time_points <- seq(0, 60, length.out = 50)
R> 
R> # Data.frame containing risk profile
R> newdata <- data.frame("Sex" = factor(c("F", "F"), 
R>                                      levels = levels(bmtcrr[,"Sex"])),
R>                       "D" = c("ALL", "AML"), # Both diseases
R>                       "Phase" = factor(c("Relapse", "Relapse"), 
R>                                        levels = levels(bmtcrr[,"Phase"])),
R>                       "Age" = c(35, 35),
R>                       "Source" = factor(c("PB", "PB"), 
R>                                         levels = levels(bmtcrr[,"Source"])))
R> 
R> # Estimate absolute risk curve
R> risk_cb <- absoluteRisk(object = model_cb, time = time_points,
R>                         method = "numerical", newdata = newdata)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(timereg)
R> model_fg <- comp.risk(Event(ftime, Status) ~ const(Sex) + const(D) + 
R>                           const(Phase) + const(Source) + const(Age), 
R>                       data = bmtcrr, cause = 1, model = "fg")
R> 
R> # Estimate absolute risk curve
R> risk_fg <- predict(model_fg, newdata, times = time_points)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> support<-casebase::support
R> 
R> support_popTime <- popTime(support,
R+                            time = "d.time", event = "death",percentile_number = 0.31)
R> plot(support_popTime)
\end{CodeInput}
\begin{figure}

@@

\begin{CodeInput}
R> completeData<-casebase::support
R> workingCompleteData=model.matrix(death~ .-d.time,data=completeData)[,-c(1)]#remove intercept
R> #Create u and xCox, which will be used when fitting Cox with glmnet.
R> newData=workingCompleteData
R> x=workingCompleteData
R> #x and y will be used to fit the casebase model under glmnet.
R> #x=as.matrix(sparse.model.matrix(death~ .-d.time+0,data=workingData))
R> y=data.matrix(completeData[,c(2,5)])
R> 
R> fullDataCox<-as.data.frame(cbind(as.numeric(y[,2]),as.numeric(y[,1]),x))
R> train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
R> test_index <- setdiff(1:nrow(fullDataCox), train_index)
R> 
R> train<-fullDataCox[train_index,]
R> test<-fullDataCox[test_index,]
R> coxSPS<- survival::coxph(Surv(time=V1,event=V2) ~ sps, data = train)
R> abcoxSPS<-survival::survfit(coxSPS,newdata = test)
R> 
R> coxAPS<- survival::coxph(Surv(time=V1,event=V2) ~ aps, data = train)
R> abcoxAPS<-survival::survfit(coxAPS,newdata = test)
R> 
R> coxFull<- survival::coxph(Surv(time=V1,event=V2) ~ .-sps -aps, data = train)
R> abcoxFull<-survival::survfit(coxAPS,newdata = test)
R> #library(prodlim)
R> 
R> 
R> KM<- survival::coxph(Surv(time=V1,event=V2) ~ 1, data = test)
R> abKM<-survival::survfit(KM)
R> #abKM <- prodlim(Hist(V1,V2)~1,data=test) minimal difference
R> abKMcompare<-cbind(abKM$time[abKM$time<=1631],1-abKM$surv[abKM$time<=1631])
R> colnames(abKMcompare)<-c("Time","KM")
R> 
R> baseCoxComparisonData<-cbind(abcoxSPS$time[abcoxSPS$time<=1631],1-rowMeans(abcoxSPS$surv[abcoxSPS$time<=1631,]),1-rowMeans(abcoxAPS$surv[abcoxSPS$time<=1631,]),1-abKM$surv,1-rowMeans(abcoxFull$surv[abcoxSPS$time<=1631,]))
R> colnames(baseCoxComparisonData)<-c("Time","SPS","APS","Unadjusted","Full")
R> myColors <- brewer.pal(5,"Dark2")
R> FigureWidth='90%'
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> ggplot(as.data.frame(ab), aes(Time)) + 
R+     geom_line(data=as.data.frame(baseCoxComparisonData),mapping=aes(y = Full, colour = "Cox"),lwd=2)+
R+   geom_line(aes(y = `Absolute risk`, colour = Model),lwd=2)+
R+   geom_step(data=as.data.frame(abKMcompare),mapping=aes(x=Time,y=KM,color="Unadjusted"),lwd=2)+
R+   scale_colour_manual(name = "Models",values = myColors[c(4,1,3,2)])+
R+   labs(y= "Probability of Death", x = "Survival time (Days)")
\end{CodeInput}
\begin{figure}

@@

\begin{CodeInput}
R> 
R> library(dotwhisker)
R> library(broom)
R> library(dplyr)
R> 
R> estimatescb=setDT(as.data.frame(coef(cbFull)[-c(1,2),1]), keep.rownames = TRUE)[]
R> #estimatescb$std.error<-0
R> estimatescb$Model="casebase"
R> colnames(estimatescb)<-c("term","estimate","model")
R> 
R> estimatescoxFull=setDT(as.data.frame(coef(coxFull)), keep.rownames = TRUE)[]
R> #estimatescoxFull$std.error<-0
R> estimatescoxFull$Model="CoxFull"
R> colnames(estimatescoxFull)<-c("term","estimate","model")
R> 
R> lolliplotDots<-rbind(estimatescb,estimatescoxNet,estimatescoxFull)
R> lolliplotDots$conf.low<-NA
R> lolliplotDots$conf.high<-NA
R> lolliplotDots$conf.high[lolliplotDots$estimate< 0]<-0
R> 
R> lolliplotDots$conf.low[lolliplotDots$estimate> 0]<-0
R> 
R> dwplot(lolliplotDots)
\end{CodeInput}

@@

\begin{CodeInput}
R> library(survival)
R> library(casebase)
R> 
R> stanford_popTime <- popTime(jasa, time = "futime", 
R>                             event = "fustat")
R> plot(stanford_popTime)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(tidyverse)
R> library(lubridate)
R> 
R> cb_data <- sampleCaseBase(jasa, time = "futime", 
R>                           event = "fustat", ratio = 10)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Define exposure variable
R> cb_data <- mutate(cb_data,
R>                   txtime = time_length(accept.dt %--% tx.date, 
R>                                        unit = "days"),
R>                   exposure = case_when(
R>                     is.na(txtime) ~ 0L,
R>                     txtime > futime ~ 0L,
R>                     txtime <= futime ~ 1L
R>                   ))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(splines)
R> # Fit several models
R> fit1 <- fitSmoothHazard(fustat ~ exposure,
R>                         data = cb_data, time = "futime")
R> fit2 <- fitSmoothHazard(fustat ~ exposure + futime,
R>                         data = cb_data, time = "futime")
R> fit3 <- fitSmoothHazard(fustat ~ exposure + bs(futime),
R>                         data = cb_data, time = "futime")
R> fit4 <- fitSmoothHazard(fustat ~ exposure*bs(futime),
R>                         data = cb_data, time = "futime")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute AIC
R> c("Model1" = AIC(fit1),
R>   "Model2" = AIC(fit2),
R>   "Model3" = AIC(fit3),
R>   "Model4" = AIC(fit4))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute hazards---
R> # First, create a list of time points for both exposure status
R> hazard_data <- expand.grid(exposure = c(0, 1),
R>                            futime = seq(0, 1000,
R>                                         length.out = 100))
R> # Set the offset to zero
R> hazard_data$offset <- 0 
R> # Use predict to get the fitted values, and exponentiate to 
R> # transform to the right scale
R> hazard_data$hazard = exp(predict(fit4, newdata = hazard_data,
R>                                  type = "link"))
R> # Add labels for plots
R> hazard_data$Status = factor(hazard_data$exposure,
R>                             labels = c("NoTrans", "Trans"))
R> 
R> ggplot(hazard_data, aes(futime, hazard, colour = Status)) +
R>     geom_line() +
R>     theme_minimal() +
R>     theme(legend.position = 'top') +
R>     ylab('Hazard') + xlab('Follow-up time')
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute absolute risk curves
R> newdata <- data.frame(exposure = c(0, 1))
R> absrisk <- absoluteRisk(fit4, newdata = newdata, 
R>                         time = seq(0, 1000, length.out = 100))
R> 
R> colnames(absrisk) <- c("Time", "NoTrans", "Trans")
R> 
R> # Rearrange the data
R> absrisk <- gather(as.data.frame(absrisk),
R>                   "Status", "Risk", -Time)
R>  
R> ggplot(absrisk, aes(Time, Risk, colour = Status)) +
R>   geom_line() +
R>   theme_minimal() +
R>   theme(legend.position = 'top') +
R>   expand_limits(y = 0.5) +
R>   xlab('Follow-up time') + ylab('Cum. Incidence')
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute hazards---
R> # First, create a list of time points for both exposure status
R> one_yr_haz <- data.frame(futime = seq(0, 365,
R>                                       length.out = 100))
R> one_yr_haz <- mutate(one_yr_haz,
R>                      offset = 0,
R>                      exposure = if_else(futime < 30, 0, 1))
R> one_yr_haz$hazard = exp(predict(fit4, newdata = one_yr_haz,
R>                                 type = "link"))
R> ggplot(one_yr_haz, aes(futime, hazard)) +
R>   geom_line() +
R>   theme_minimal() +
R>   theme(legend.position = 'top') +
R>   ylab('Hazard') + xlab('Follow-up time') +
R>   geom_vline(xintercept = 30, linetype = 'dashed')
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> absoluteRisk(fit4, newdata = data.frame(exposure = 0),
R>              time = 365)
R> 
R> # Use the trapezoidal rule to estimate the cumulative hazard
R> cumhaz_est <- pracma::trapz(one_yr_haz$futime, one_yr_haz$hazard)
R> 1 - exp(-cumhaz_est)
\end{CodeInput}
\end{CodeChunk}

@@

This report was generated on 2020-03-05 23:20:03 using the following
computational environment and dependencies:

>>>>

\begin{CodeInput}
R> data(ERSPC)
R> ERSPC$ScrArm <- factor(ERSPC$ScrArm,
R>   levels = c(0, 1),
R>   labels = c("Control group", "Screening group")
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> pt_object_strat <- casebase::popTime(ERSPC,
R>   event = "DeadOfPrCa",
R>   exposure = "ScrArm"
R> )
R> plot(pt_object_strat)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> cox_model <- survival::coxph(Surv(Follow.Up.Time, DeadOfPrCa) ~ ScrArm,
R>   data = ERSPC
R> )
R> summary(cox_model)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> new_data <- data.frame(
R>   ScrArm = c("Control group", "Screening group"),
R>   ignore = 99
R> )
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>   xlab = "Years since Randomization",
R>   ylab = "Cumulative Incidence (%)",
R>   fun = "event",
R>   xlim = c(0, 15), conf.int = FALSE, col = c("red", "blue"),
R>   main = sprintf(
R>     "Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>     exp(coef(cox_model)),
R>     exp(confint(cox_model))[1],
R>     exp(confint(cox_model))[2]
R>   ),
R>   ylim = c(0, 2), yscale = 100
R> )
R> legend("topleft",
R>   legend = c("Control group", "Screening group"),
R>   col = c("red", "blue"),
R>   lty = c(1, 1),
R>   bg = "gray90"
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> casebase_exponential <- casebase::fitSmoothHazard(DeadOfPrCa ~ ScrArm,
R>   data = ERSPC,
R>   ratio = 100
R> )
R> 
R> summary(casebase_exponential)
R> exp(coef(casebase_exponential)[2])
R> exp(confint(casebase_exponential)[2, ])
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> smooth_risk_exp <- casebase::absoluteRisk(
R>   object = casebase_exponential,
R>   time = seq(0, 15, 0.1),
R>   newdata = new_data
R> )
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>   xlab = "Years since Randomization",
R>   ylab = "Cumulative Incidence (%)",
R>   fun = "event",
R>   xlim = c(0, 15), conf.int = FALSE, col = c("red", "blue"),
R>   main = sprintf(
R>     "Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>     exp(coef(cox_model)),
R>     exp(confint(cox_model))[1],
R>     exp(confint(cox_model))[2]
R>   ),
R>   ylim = c(0, 2), yscale = 100
R> )
R> lines(smooth_risk_exp[, 1], smooth_risk_exp[, 2], col = "red", lty = 2)
R> lines(smooth_risk_exp[, 1], smooth_risk_exp[, 3], col = "blue", lty = 2)
R> 
R> 
R> legend("topleft",
R>   legend = c(
R>     "Control group (Cox)", "Control group (Casebase)",
R>     "Screening group (Cox)", "Screening group (Casebase)"
R>   ),
R>   col = c("red", "red", "blue", "blue"),
R>   lty = c(1, 2, 1, 2),
R>   bg = "gray90"
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> casebase_time <- fitSmoothHazard(DeadOfPrCa ~ Follow.Up.Time + ScrArm,
R>   data = ERSPC,
R>   ratio = 100
R> )
R> 
R> summary(casebase_time)
R> exp(coef(casebase_time))
R> exp(confint(casebase_time))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> smooth_risk_time <- casebase::absoluteRisk(
R>   object = casebase_time,
R>   time = seq(0, 15, 0.1),
R>   newdata = new_data
R> )
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>   xlab = "Years since Randomization",
R>   ylab = "Cumulative Incidence (%)",
R>   fun = "event",
R>   xlim = c(0, 15), conf.int = FALSE, col = c("red", "blue"),
R>   main = sprintf(
R>     "Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>     exp(coef(cox_model)),
R>     exp(confint(cox_model))[1],
R>     exp(confint(cox_model))[2]
R>   ),
R>   ylim = c(0, 2), yscale = 100
R> )
R> lines(smooth_risk_time[, 1], smooth_risk_time[, 2], col = "red", lty = 2)
R> lines(smooth_risk_time[, 1], smooth_risk_time[, 3], col = "blue", lty = 2)
R> 
R> legend("topleft",
R>   legend = c(
R>     "Control group (Cox)", "Control group (Casebase)",
R>     "Screening group (Cox)", "Screening group (Casebase)"
R>   ),
R>   col = c("red", "red", "blue", "blue"),
R>   lty = c(1, 2, 1, 2),
R>   bg = "gray90"
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> casebase_splines <- fitSmoothHazard(DeadOfPrCa ~ bs(Follow.Up.Time) + ScrArm,
R>   data = ERSPC,
R>   ratio = 100
R> )
R> 
R> summary(casebase_splines)
R> exp(coef(casebase_splines))
R> exp(confint(casebase_splines))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> smooth_risk_splines <- absoluteRisk(
R>   object = casebase_splines,
R>   time = seq(0, 15, 0.1),
R>   newdata = new_data
R> )
R> 
R> plot(survfit(cox_model, newdata = new_data),
R>   xlab = "Years since Randomization",
R>   ylab = "Cumulative Incidence (%)",
R>   fun = "event",
R>   xlim = c(0, 15), conf.int = FALSE, col = c("red", "blue"),
R>   main = sprintf(
R>     "Estimated Cumulative Incidence (risk) of Death from Prostate 
R>                     Cancer Screening group Hazard Ratio: %.2g (%.2g, %.2g)",
R>     exp(coef(cox_model)),
R>     exp(confint(cox_model))[1],
R>     exp(confint(cox_model))[2]
R>   ),
R>   ylim = c(0, 2), yscale = 100
R> )
R> lines(smooth_risk_splines[, 1], smooth_risk_splines[, 2], col = "red", lty = 2)
R> lines(smooth_risk_splines[, 1], smooth_risk_splines[, 3], col = "blue", lty = 2)
R> 
R> legend("topleft",
R>   legend = c(
R>     "Control group (Cox)", "Control group (Casebase)",
R>     "Screening group (Cox)", "Screening group (Casebase)"
R>   ),
R>   col = c("red", "red", "blue", "blue"),
R>   lty = c(1, 2, 1, 2),
R>   bg = "gray90"
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> c(
R>   "Exp." = AIC(casebase_exponential),
R>   "Weib." = AIC(casebase_time),
R>   "Splines" = AIC(casebase_splines)
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> model_cb <- fitSmoothHazard(
R>   Status ~ ftime + Sex + D + Phase + Source + Age,
R>   data = bmtcrr,
R>   ratio = 100,
R>   time = "ftime"
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(survival)
R> # Treat competing event as censoring
R> model_cox <- coxph(Surv(ftime, Status == 1) ~ Sex + D + Phase + Source + Age,
R>   data = bmtcrr
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Pick 100 equidistant points between 0 and 60 months
R> time_points <- seq(0, 60, length.out = 50)
R> 
R> # Data.frame containing risk profile
R> newdata <- data.frame(
R>   "Sex" = factor(c("F", "F"),
R>     levels = levels(bmtcrr[, "Sex"])
R>   ),
R>   "D" = c("ALL", "AML"), # Both diseases
R>   "Phase" = factor(c("Relapse", "Relapse"),
R>     levels = levels(bmtcrr[, "Phase"])
R>   ),
R>   "Age" = c(35, 35),
R>   "Source" = factor(c("PB", "PB"),
R>     levels = levels(bmtcrr[, "Source"])
R>   )
R> )
R> 
R> # Estimate absolute risk curve
R> risk_cb <- absoluteRisk(
R>   object = model_cb, time = time_points,
R>   method = "numerical", newdata = newdata
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(timereg)
R> model_fg <- comp.risk(Event(ftime, Status) ~ const(Sex) + const(D) +
R>   const(Phase) + const(Source) + const(Age),
R> data = bmtcrr, cause = 1, model = "fg"
R> )
R> 
R> # Estimate absolute risk curve
R> risk_fg <- predict(model_fg, newdata, times = time_points)
\end{CodeInput}
\end{CodeChunk}

@@

@@

@@

\begin{CodeInput}
R> 
R> cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = ratio, standardize = TRUE)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeInput}
R> # Estimating the absolute risk curve using the newData parameter.
R> abCbFull <- casebase::absoluteRisk(cbFull, time = seq(1, max(abKMcompare[, 1]), 25), newdata = newData, s = "lambda.1se", method = c("numerical"))
\end{CodeInput}
\end{CodeChunk}

\begin{CodeInput}
R> ggplot(as.data.frame(ab), aes(Time)) +
R+   geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 2) +
R+   geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 2) +
R+   geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 2) +
R+   labs(y = "Probability of Death", x = "Survival time (Days)")
\end{CodeInput}
\begin{figure}

@@

\begin{CodeInput}
R> 
R> library(dotwhisker)
R> library(broom)
R> library(dplyr)
R> 
R> estimatescb <- setDT(as.data.frame(coef(cbFull)[-c(1, 2), 1]), keep.rownames = TRUE)[]
R> # estimatescb$std.error<-0
R> estimatescb$Model <- "casebase"
R> colnames(estimatescb) <- c("term", "estimate", "model")
R> 
R> estimatescoxFull <- setDT(as.data.frame(coef(coxFull)), keep.rownames = TRUE)[]
R> # estimatescoxFull$std.error<-0
R> estimatescoxFull$Model <- "CoxFull"
R> colnames(estimatescoxFull) <- c("term", "estimate", "model")
R> 
R> lolliplotDots <- rbind(estimatescb, estimatescoxNet, estimatescoxFull)
R> lolliplotDots$conf.low <- lolliplotDots$estimate
R> lolliplotDots$conf.high <- lolliplotDots$estimate
R> lolliplotDots$conf.high[lolliplotDots$estimate < 0] <- 0
R> lolliplotDots$conf.low[lolliplotDots$estimate > 0] <- 0
R> 
R> dwplot(lolliplotDots)
\end{CodeInput}

@@

\begin{CodeInput}
R> library(survival)
R> library(casebase)
R> 
R> stanford_popTime <- popTime(jasa,
R>   time = "futime",
R>   event = "fustat"
R> )
R> plot(stanford_popTime)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(tidyverse)
R> library(lubridate)
R> 
R> cb_data <- sampleCaseBase(jasa,
R>   time = "futime",
R>   event = "fustat", ratio = 10
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Define exposure variable
R> cb_data <- mutate(cb_data,
R>   txtime = time_length(accept.dt %--% tx.date,
R>     unit = "days"
R>   ),
R>   exposure = case_when(
R>     is.na(txtime) ~ 0L,
R>     txtime > futime ~ 0L,
R>     txtime <= futime ~ 1L
R>   )
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> library(splines)
R> # Fit several models
R> fit1 <- fitSmoothHazard(fustat ~ exposure,
R>   data = cb_data, time = "futime"
R> )
R> fit2 <- fitSmoothHazard(fustat ~ exposure + futime,
R>   data = cb_data, time = "futime"
R> )
R> fit3 <- fitSmoothHazard(fustat ~ exposure + bs(futime),
R>   data = cb_data, time = "futime"
R> )
R> fit4 <- fitSmoothHazard(fustat ~ exposure * bs(futime),
R>   data = cb_data, time = "futime"
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute AIC
R> c(
R>   "Model1" = AIC(fit1),
R>   "Model2" = AIC(fit2),
R>   "Model3" = AIC(fit3),
R>   "Model4" = AIC(fit4)
R> )
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute hazards---
R> # First, create a list of time points for both exposure status
R> hazard_data <- expand.grid(
R>   exposure = c(0, 1),
R>   futime = seq(0, 1000,
R>     length.out = 100
R>   )
R> )
R> # Set the offset to zero
R> hazard_data$offset <- 0
R> # Use predict to get the fitted values, and exponentiate to
R> # transform to the right scale
R> hazard_data$hazard <- exp(predict(fit4,
R>   newdata = hazard_data,
R>   type = "link"
R> ))
R> # Add labels for plots
R> hazard_data$Status <- factor(hazard_data$exposure,
R>   labels = c("NoTrans", "Trans")
R> )
R> 
R> ggplot(hazard_data, aes(futime, hazard, colour = Status)) +
R>   geom_line() +
R>   theme_minimal() +
R>   theme(legend.position = "top") +
R>   ylab("Hazard") +
R>   xlab("Follow-up time")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute absolute risk curves
R> newdata <- data.frame(exposure = c(0, 1))
R> absrisk <- absoluteRisk(fit4,
R>   newdata = newdata,
R>   time = seq(0, 1000, length.out = 100)
R> )
R> 
R> colnames(absrisk) <- c("Time", "NoTrans", "Trans")
R> 
R> # Rearrange the data
R> absrisk <- gather(
R>   as.data.frame(absrisk),
R>   "Status", "Risk", -Time
R> )
R> 
R> ggplot(absrisk, aes(Time, Risk, colour = Status)) +
R>   geom_line() +
R>   theme_minimal() +
R>   theme(legend.position = "top") +
R>   expand_limits(y = 0.5) +
R>   xlab("Follow-up time") +
R>   ylab("Cum. Incidence")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Compute hazards---
R> # First, create a list of time points for both exposure status
R> one_yr_haz <- data.frame(futime = seq(0, 365,
R>   length.out = 100
R> ))
R> one_yr_haz <- mutate(one_yr_haz,
R>   offset = 0,
R>   exposure = if_else(futime < 30, 0, 1)
R> )
R> one_yr_haz$hazard <- exp(predict(fit4,
R>   newdata = one_yr_haz,
R>   type = "link"
R> ))
R> ggplot(one_yr_haz, aes(futime, hazard)) +
R>   geom_line() +
R>   theme_minimal() +
R>   theme(legend.position = "top") +
R>   ylab("Hazard") +
R>   xlab("Follow-up time") +
R>   geom_vline(xintercept = 30, linetype = "dashed")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> absoluteRisk(fit4,
R>   newdata = data.frame(exposure = 0),
R>   time = 365
R> )
R> 
R> # Use the trapezoidal rule to estimate the cumulative hazard
R> cumhaz_est <- pracma::trapz(one_yr_haz$futime, one_yr_haz$hazard)
R> 1 - exp(-cumhaz_est)
\end{CodeInput}
\end{CodeChunk}

@@

This report was generated on 2020-03-06 08:01:48 using the following
computational environment and dependencies:
